White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-09-09 papers/wdmerger_II/paper.tex
Re-title paper II, rework the abstract

%==========================================================================
% Title
%==========================================================================
\title{White Dwarf Mergers on Adaptive Meshes\\ II. Collisions}

\shorttitle{WD Mergers. II. Collisions}
\shortauthors{Katz et al. (2017)}

@@

%==========================================================================
% Abstract
%==========================================================================
\begin{abstract}
Collisions of white dwarfs are a subset of white dwarf mergers that occur
when the stars approach each other nearly head-on. These have the potential
for explosive nucleosynthesis, leading to a thermonuclear detonation that
disrupts both stars, possibly causing a Type Ia supernova. Simulation of
these events is challenging due to the rapid evolution of the nuclear burning,
posing numerical challenges related to both spatial and temporal resolution.
In this study we explore a number of the algorithms that affect the outcome
of reacting hydrodynamics simulations that lead to explosive burning and
address the question of what needs to be done to ensure that these simulations
are properly converged. We find that achieving a converged simulation in the
code \castro\ is non-trivial and requires demanding computational resources.
We also discuss the quantities obtained from our simulations that bear relevance
to observations of these events, particularly production of nickel-group
elements and gravitational wave emission during the collision.
\end{abstract}
\keywords{supernovae: general - white dwarfs}

>>>>

%==========================================================================
% Title
%==========================================================================
\title{White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions}

\shorttitle{WD Mergers. II. Nuclear Reactions}
\shortauthors{Katz et al. (2017)}

@@

%==========================================================================
% Abstract
%==========================================================================
\begin{abstract}
Simulations of white dwarf mergers and collisions that attempt to answer the
question of whether these events are progenitors of Type Ia supernovae must
confront the question of how nuclear detonations form and whether observed
detonations in a numerical simulation are realistic. We address this question
by performing simulations of collisions of white dwarfs, a subset of white
dwarf mergers that occur when the stars approach each other nearly head-on.
These have the potential for explosive nucleosynthesis, leading to a thermonuclear
detonation that disrupts both stars, possibly causing a Type Ia supernova.
While this story is intuitively plausible, demonstrating this in a reacting
hydrodynamics simulation is challenging due to the rapid evolution of the
nuclear burning, posing numerical challenges related to both spatial and
temporal resolution. In this study we explore a number of the algorithms
that affect the outcome of reacting hydrodynamics simulations that lead
to explosive burning and address the question of what can be done to ensure
that these simulations are properly converged. We find that achieving a
converged simulation in the code \castro\ is non-trivial, and it is not
clear that we can correctly simulate a white dwarf collision and obtain
an adequate estimate of the nucleosynthetic yield. This result calls into
question whether it is possible to self-consistently simulate a white dwarf
merger leading to a thermonuclear detonation.
\end{abstract}
\keywords{supernovae: general - white dwarfs}
